---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# electrocpue

<!-- badges: start -->
[![R-CMD-check](https://github.com/shepherd70/electrocpue/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/shepherd70/electrocpue/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`electrocpue` turns raw multi-pass electrofishing records into
abundance, density, and catch-per-unit-effort (CPUE) estimates. It
provides assumption-aware removal-depletion population estimators
(Zippin and Carle & Strub), effort standardization (raw seconds or
amp-seconds), and tidy, plot-ready output.

Point estimates and standard errors match `FSA::removal()`, so results
are directly comparable to that reference implementation — but FSA is
only a test-time dependency, not required at runtime.

## Installation

You can install the development version of electrocpue from
[GitHub](https://github.com/shepherd70/electrocpue) with:

``` r
# install.packages("pak")
pak::pak("shepherd70/electrocpue")
```

## Workflow

The package follows a four-step workflow: **validate → estimate →
analyze → summarize**.

```{r library}
library(electrocpue)
```

### Estimate a single removal series

`estimate_population()` works on a vector of per-pass catches. The
default `"auto"` method uses the Zippin maximum-likelihood estimator and
falls back to Carle & Strub when the Zippin model fails.

```{r estimate}
estimate_population(c(45, 18, 7))
```

### Analyze a whole dataset

`analyze_cpue()` validates the inputs, estimates abundance for every
reach × date × species series, standardizes effort, and converts
abundance to density. It returns one tidy row per series.

```{r analyze}
res <- analyze_cpue(example_catch, example_reach)
head(res[, c("reach_id", "date", "species", "N", "N_se", "density_per_m", "cpue")])
```

### Summarize repeat surveys

`summarize_cpue()` rolls repeat survey dates up to a reach × species
summary with confidence intervals on density.

```{r summarize}
summarize_cpue(res)[, c("reach_id", "species", "n_surveys", "N_mean",
                        "density_per_m_mean", "density_per_m_lwr",
                        "density_per_m_upr")]
```

## Learn more

See `vignette("electrocpue")` for a complete worked example, including
effort standardization by amp-seconds and plotting density with
confidence intervals.
