---
title: "Analyzing electrofishing CPUE with electrocpue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing electrofishing CPUE with electrocpue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 96
)
```

## Overview

`electrocpue` turns raw multi-pass electrofishing records into
abundance, density, and catch-per-unit-effort (CPUE) estimates. The
workflow has four steps:

1. **Validate** the catch and reach-metadata tables.
2. **Estimate** removal-depletion abundance for each removal series.
3. **Analyze** — wire validation, estimation, effort standardization,
   and density together in one call.
4. **Summarize** repeat surveys up to a reach level with confidence
   intervals.

```{r library}
library(electrocpue)
```

## The example data

The package ships a small simulated dataset: four reaches, each surveyed
on two dates, with two species and three removal passes per survey.

```{r data}
head(example_catch)

example_reach
```

`example_catch` is **long format**: one row per reach × date × pass ×
species. `example_reach` carries the reach length (and, here, area) used
to convert abundance into density.

## Step 1 — Validate

`validate_cpue_input()` runs a battery of structural and content checks
and reports every problem at once. It returns invisibly on success.

```{r validate}
validate_cpue_input(example_catch, example_reach, strict = FALSE)
```

## Step 2 — Estimate a single series

The removal estimators work on a vector of per-pass catches. `"auto"`
uses the Zippin maximum-likelihood estimator and falls back to Carle &
Strub when the Zippin model fails.

```{r estimate}
# Brown trout at R01, first visit: 82, 42, 19
estimate_population(c(82, 42, 19), method = "auto")
```

The result reports the abundance estimate `N`, its standard error
`N_se`, the per-pass capture probability `p`, and a `converged` flag.

## Step 3 — Analyze the whole dataset

`analyze_cpue()` does everything at once and returns one tidy row per
reach × date × species. Here we standardize effort by amp-seconds, which
requires the `amperage` column present in `example_catch`.

```{r analyze}
res <- analyze_cpue(
  example_catch, example_reach,
  method       = "auto",
  effort_basis = "amp_seconds"
)

res[, c("reach_id", "date", "species", "N", "N_se",
        "density_per_m", "density_per_m2", "cpue")]
```

Both length-based (`density_per_m`) and areal (`density_per_m2`) density
are returned; areal density is `NA` when reach area is unavailable.

## Step 4 — Summarize repeat surveys

`summarize_cpue()` collapses the repeat visits per reach to a single
reach × species summary with a confidence interval on density. The
interval pools each survey's profile-likelihood uncertainty with the
between-survey variation on the log scale. A reach is flagged `weak`
(and `n_identified` records how many surveys backed the interval) when
too few surveys are well enough identified to support it — at low capture
probability the removal estimate is biased, so the interval is withheld
rather than shown as a confident bar.

```{r summarize}
smry <- summarize_cpue(res, level = 0.95)
smry[, c("reach_id", "species", "n_surveys", "n_identified", "weak",
         "N_mean", "density_per_m_mean",
         "density_per_m_lwr", "density_per_m_upr")]
```

## Visualizing density with uncertainty

Identified reaches carry a 95% interval; a `weak` reach is drawn as a
hollow point with no bar (the interval would not be trustworthy).

```{r plot, eval = requireNamespace("ggplot2", quietly = TRUE)}
library(ggplot2)

ggplot(smry, aes(x = reach_id, y = density_per_m_mean, colour = species)) +
  geom_errorbar(
    aes(ymin = density_per_m_lwr, ymax = density_per_m_upr),
    width = 0.2,
    position = position_dodge(width = 0.4)
  ) +
  geom_point(aes(shape = weak),
             position = position_dodge(width = 0.4), size = 2.5) +
  scale_shape_manual(
    values = c(`FALSE` = 16, `TRUE` = 1),
    labels = c(`FALSE` = "identified", `TRUE` = "weak"),
    name   = "interval"
  ) +
  labs(
    x = "Reach", y = "Density (fish / m)", colour = "Species",
    title = "Estimated linear density by reach",
    subtitle = "Reach mean; 95% profile-likelihood pooled intervals"
  ) +
  theme_minimal(base_size = 12)
```

## Choosing an estimator

- **`"zippin"`** — classic maximum-likelihood depletion estimator;
  reports `NA` when the catch series shows insufficient depletion.
- **`"carle_strub"`** — weighted-likelihood (Bayesian) estimator; more
  robust when depletion is weak, and the recommended default for routine
  use.
- **`"auto"`** — Zippin first, Carle & Strub as a fallback.

Point estimates and standard errors match `FSA::removal()`, so results
are directly comparable to that reference implementation.
